---
title: "Predict Price"
author: "Hieu Trung Nguyen"
date: "8/1/2022"
output: html_document
---
In this project, we will predict Price variable from dataset Computers in the Ecdat package. 
a. Record categorical variables as factors and remove variables that cannot be used in the analysis. 

```{r}
# loading library. In case there is no library install yet, we can use the code: 
# install.packages('Ecdat')
library(Ecdat)

data(Computers) # calling for the dataset
# View(Computers)
len <- sapply(Computers, table) # display the length of each varaible in the Computers dataset
# for the variables that having less than 5 values, we change them as factor
Computers$screen <- as.factor(Computers$screen)
Computers$cd <- as.factor(Computers$cd)
Computers$multi <- as.factor(Computers$multi)
Computers$premium <- as.factor(Computers$premium)
sapply(Computers, class) # check the class of varaible in Computers
```
```{r}
n <- nrow(Computers) # counting the rows of Computers and assign it to n
ntrain <- floor(n/2) # having the ntrain variable as 50% of the n variable
set.seed(4598) # set seed as 4598
ind <- sample(1:n, ntrain) # having ind as randomly sample of all n, size of ind = ntrain
train <- Computers[ind, ] # having all variables value (observations) that we sampling from ind
test <- Computers[-ind, ]
```
b. use descriptive statistic to find promising predictors 
```{r}
plot(price~., data = train)
```


c. run a linear regression on all explanatory variables. Interpret some of the coefficients. Evaluate the prediction performance on an appropriate measure. 
```{r}
reg1 <- lm(price~., data = train)  # run the linear regression of of price to all other varaibles.
summary(reg1)
```
All variables are significant.
```{r}

pred1 <- predict(reg1, newdata = test) # predict price from the result reg1 using new dataset test
MSE1 <- mean((test$price-pred1)^2) # calcualte mean square error
MSE1

```

d. investigate if the predictions of price would be better if we use a linear regression of log(price) on the other variables.
```{r}
Computers$logp <- log(Computers$price) # create new variable logp by using log(price)
train <- Computers[ind, ] # create new train dataset including new variable logp
test <- Computers[-ind,]
reg2 <- lm(logp~.-price, data = train) # using linear regression using all other variables except price to predict price
pred2 <- exp(predict(reg2, newdata=test)) # predict logp from the test dataset using the result from reg2
MSE2 <- mean((test$price-pred2)^2)  # calculate the mean square error
MSE2
MSE1 - MSE2
```
With logp, the MSE is higher than without log. MSE the lower the value the better and 0 means the model is perfect.
e. in d we fitted a model where price is a particular nonlinear function of the other variables. We now investigate other non-linear models. First, fit a GAM-model, plot the result and evalute the predictions.
```{r}
# call for library gam
library(gam)
gam1 <- gam(price~s(speed) + s(hd) + s(ram) + screen + cd + multi + premium + s(ads) + s(trend), data=train) # call gam prediction, for numeric variable. s() Function used in definition of smooth terms within gam model formulae. The function does not evaluate a (spline) smooth - it exists purely to help set up a model using spline based smooths.
par(mfrow=c(3,3)) # set the plot for 3*3 images 
plot(gam1) # plot the gam1 result
```
```{r}
pred3 <- predict(gam1, newdata = test) # predict price using newdata test, based on the gam1 result
MSE3 <- mean((test$price-pred3)^2)
MSE3
MSE1 - MSE3
```
GAM improved the prediction quite a lot

f. give a brief explantion of how backfitting works and in what situation it is possible to fit a generalized additive model with ordinary least square regression. 
In backfitting a model of the form: 
  y = f1(x1) + f2(x2) +  ϵ 
is estimated by first setting f2(x2) = 0 and estimate f1 with a method for one-predictor models, e.g., the Nadaraya-Watson estimator. The residuals e1 = y -f1^(x1) are then computed and will substitute the response variable in the next iteration where e is fitted to f2(x2). Now, the residuals e2 = e1 - f2^(x2) are computed and again fitted to f1(x1). this procedure is continued until the estimators of f1 and f2 converges. 
A GAM can be fitted with OLS if the functions fk(xk) can be written with so called basis functions. Examples of that are splines and piecewise polynomials. 

g. Use bagged trees to predict price. Evaluate the predictions. Compute varaible importance measures and comment on the results. Explain the logic behind the measure for variable importance using.
```{r}
# loading library 
library(randomForest)

bag1 <- randomForest(price~. - logp, data=train, mtry=9, ntree=100) # predict price using all variables execpt logp of the train dataset. Number of randomly sampled predictors = 9, the number of bootstrap replicates = 100
pred4 <- predict(bag1, newdata = test) # predict price using test dataset based on the result of bag1
MSE4 <- mean((test$price-pred4)^2)
MSE4
MSE3-MSE4
```
The prediction result is significantly improved. 
```{r}
varImpPlot(bag1)
```


The measure used is average decrease in MSE over all trees obtained when a variable is added. 

h. Give a brief explanation of bagging 
In bagging, bootstrap samples (a draw, with replacement, of the same size as original sample) are use to fit a model, e.g., a regression tree. A prediction is done by the model for each bootstrap sample and the final prediction is computed by averaging all of them.
